---
title: "CS-201 First Look at R"
author: "Amber Camp"
date: 8/26/26
format: html
editor: visual
toc: true
---

This is the first class session walking through code together!

We will be programming in R and in this class will primarily use `qmd` documents to keep things tidy and presentable. Each code chunk below is where you will run R code. Feel free to add notes as we go along.

## Navigating RStudio

By default, everyone should have the same four-pane layout:

1.  **Top left**: Source editor

2.  **Bottom left**: Console (and Terminal and Background Jobs)

3.  **Top Right**: Environment

4.  **Bottom Right**: Files, Plots, Help, etc.

### Customizing look & feel

Feel free to customize your RStudio.\
Go to Tools \> Global Options \> Appearance to select a color scheme you like.

I also personally like to enable these:

- Tools \> Global Options \> Code \> Insert matching parens/quotes

- Tools \> Global Options \> Console \> Show syntax highlighting in console input

I don't recommend rearranging your pane layout at this time.

## Running Code

Try running some code

### Load Library

```{r}
library(ggplot2)
```

### Some Scripting

```{r}
# This is a comment
# R runs line by line, or highlighted text

quiz1 <- 88 # how to assign values to names
quiz2 <- 94
quiz3 <- 79

quiz1

# averaging
(quiz1 + quiz2 +quiz3) / 3

mean(quiz1 + quiz2 + quiz3) # Not correct. Adds them together
mean(quiz1, quiz2, quiz3) # Not correct. Only returns mean of first argument.

mean(c(quiz1, quiz2, quiz3)) # Correct

list <- c(quiz1, quiz2, quiz3) # Assign multiple values to one list

mean(list)

average <- mean(list)

quiz3 <- 99 # Can update values

mean(c(quiz1, quiz2, quiz3))
list <- c(quiz1, quiz2, quiz3)

average <- mean(list)

average
```

### More Scripts

```{r}
scores <- c(88, 94, 79, 91, 85)
scores

mean(scores)
min(scores) # smallest number
max(scores) # largest number
length(scores) # size of list

summary(scores) # all those in one keyword

students <- c("a", "b", "c", "d", "e")
students

# bar plot

barplot(scores, names.arg = students)
```

## Introducing `ggplot2`

```{r}


```

## Extra Credit (1pt)

What result do you expect when you run this expression?

```{r}

```

What did you observe?

> Answer:

Why do you think this is happening? (You can look online for help)

> Answer:
