---
title: "Assignment 4: Exploratory Data Analysis"
author: "Your Name"
date: "`r Sys.Date()`"
format: html
editor: visual
---

## Overview

Due: **Friday 9/25 11:59pm**

In this assignment, you will practice Exploratory Data Analysis (EDA) by making plots from a **Kaggle dataset of your choice**.

You will submit **at least 4 different plot types**, with short written interpretations for each.

Part 1 is worth a total of 1 point, and Part 2 is worth 4 points (1 point for each plot, question, and interpretation).

## Part 1: Get Your Data

1.  Go to [Kaggle Datasets](https://www.kaggle.com/datasets), create an account, and choose a dataset that interests you. Ensure your data has:
    1.  At least 4 variables
    2.  At least one categorical and one continuous variable
    3.  I recommend data that contains at least 50-100 observations
    4.  Tip: Open the CSV on Kaggle first and skim the columns to check that it meets these requirements
2.  Download the dataset as a CSV file.
3.  Place the CSV in your class R project `data` folder.
4.  Read it into R with `read_csv()`.

```{r, message=FALSE}
library(tidyverse)
library(ggplot2)
library(plotly)

five_hundred_hits <- read_csv("../data/500hits.csv")

five_hundred_hits$PLAYER[151] <- "Ivan Rodriguez"
# This must be done because his name in the unedited dataset has Spanish tildes, which R has a hard time parsing, so I changed his name to remove the tildes.

five_hundred_hits <- five_hundred_hits |> 
  mutate(HOF = recode(as.character(HOF),
                          "0" = "No",
                          "1" = "Yes",
                          .missing = "unknown"),
         HOF = factor(HOF))
# changes the stat for whether the player is in the Hall of Fame from a 1 or 0 into a Yes or No factor

five_hundred_hits <- five_hundred_hits |> 
  mutate(across(where(is.numeric), ~na_if(.x, 0)))
# For some reason, instead of labeling it as such, NA values are signified by 0, so I changed them to NA in order to make them easier to omit from visualizations.
```

### Describe your data

Data author: Krupa_20, or krupadharamshi

Link: [https://www.kaggle.com/datasets/krupadharamshi/500hits](https://www.kaggle.com/datasets/krupadharamshi/500hits?select=500hits.csv)

Description: A database containing a diverse array of various baseball players, and their essential in-game hitting statistics, like hits, games played, at-bats, and more.

### Preview your data

Use some of the strategies you've learned in this class or picked up elsewhere to summarize and quickly explore your data.

```{r}

head(five_hundred_hits)
```

### Summary stats

Before you make plots, you must also:

- Show basic summaries of your dataset, including:

  - `count()` for categorical variables (to see category frequencies)

  - `mean()`, `median()`, `sd()`, `min()`, `max()` for at least one numeric variable

  - Remember to use `na.rm = TRUE` if you have missing values that affect summaries!

```{r}

summary(five_hundred_hits)

count(five_hundred_hits, HOF)

mean(five_hundred_hits$H, na.rm = TRUE)

sd(five_hundred_hits$HR, na.rm = TRUE)
```

Write 2–3 sentences here: What do the category counts tell you? What does the numeric summary show about spread, center, or outliers?

The category count tells me that the majority of the players in this dataset did not make it into the Hall of Fame, around 75% did not make it in. The numeric summary also shows me that the ranges for many of these stats are large. However, I do notice a very large difference between the medians and the maximums in many columns, which tells me that there are lots of outliers, particularly to upper part of the dataset.

## Part 2: Visualize Your Data

Create **at least 4 plots**, each of a different type. Your plots may share variables, but each one should answer a **different question** about your data. At least one plot must include a variable that isn't in the others.

Each plot should have:

- A title and labeled axes

- A short interpretation (2-3 sentences) underneath

  - Sample interpretation: (boxplot of `price` by `cut` from `diamonds` data): *Ideal diamonds have the lowest median price, which is surprising since Ideal is the best cut. Every cut also has a lot of very expensive outliers. One possible reason for the low Ideal prices is that Ideal diamonds might be smaller on average, which I could check by plotting carat by cut.*

Use the EDA Part 3 notes as your reference. Since this is an EDA assignment, your plots can look plain, but consider adding colors, themes, or interactivity for interpretability and visual appeal if you know how.

### Plot 1

```{r}
# insert code here

ggplot(five_hundred_hits, aes(x = HOF, y = BB)) +
  geom_boxplot(na.rm = TRUE) +
  labs(title = "Range of Walks", x = "Hall of Fame status", y = "Total Walks")

```

Question this plot answers: What is the range of values for Base on Balls (Walks) by Hall of Fame status in the five_hundred_hits dataframe?

Interpretation: Players who have been inducted into the Hall of Fame tended to have higher walk rates, with considerable outliers. This makes sense, as the best hitters are often the ones who are able to get on base, regardless if they got a hit or not. A great hitter is able to take advantage of a pitcher who is missing the strike zone, as well as being able to stay disciplined and not swing at pitches outside the zone.

### Plot 2

```{r}
# insert code here

ggplot(five_hundred_hits, aes(x = YRS)) +
  geom_density() +
  labs(title = "Density of Years Played", x = "Amount of Years", y = "Density") +
  scale_x_continuous(breaks = seq(10, 30, by = 2))

```

Question this plot answers: What is the distribution of total years played in the MLB by all players in the five_hundred_hits dataframe?

Interpretation: The most common amount of total years a player spends in the MLB appears to be around 17 years. The drop-off occurring around Year 18 and beyond suggests that it's not common to be a long-term player in the MLB, which makes sense as the body will probably start to degrade at around this time, and only the best are able to keep playing past their physical prime and still perform well.

### Plot 3

```{r}
# insert code here

ggplot(five_hundred_hits, aes(x = BA, y = R)) +
  geom_point(na.rm = TRUE) +
  geom_smooth(method = "lm")


```

Question this plot answers: Is there a correlation between having a high batting average and scoring lots of runs in the five_hundred_hits dataframe?

Interpretation: There appears to be a light, but not very strong correlation between scoring runs and having a high batting average. This is surprising, considering often times a base hit is what drives in a run, and that a home run (even a solo home run) is counted as both an RBI and a hit. There are ways to score a runner without a base hit, such as as bases-loaded walk, a sacrifice fly, or maybe they were thrown out but the run was still able to score. As a note, the only way a run scored is NOT counted as an RBI is when the runners scores on a fielding error, on a groundball that results in a double play.

### Plot 4

```{r}
# insert code here


```

Question this plot answers:

Interpretation:

------------------------------------------------------------------------

# Push to GitHub AND submit link on Canvas!!

Assignment is due Friday 9/25 at 11:59 pm.
