---
title: "CS-201 First Look at R"
author: "Amber Camp"
date: 8/26/26
format: html
editor: visual
toc: true
---

This is the first class session walking through code together!

We will be programming in R and in this class will primarily use `qmd` documents to keep things tidy and presentable. Each code chunk below is where you will run R code. Feel free to add notes as we go along.

## Navigating RStudio

By default, everyone should have the same four-pane layout:

1.  **Top left**: Source editor

2.  **Bottom left**: Console (and Terminal and Background Jobs)

3.  **Top Right**: Environment

4.  **Bottom Right**: Files, Plots, Help, etc.

### Customizing look & feel

Feel free to customize your RStudio.\
Go to Tools \> Global Options \> Appearance to select a color scheme you like.

I also personally like to enable these:

- Tools \> Global Options \> Code \> Insert matching parens/quotes

- Tools \> Global Options \> Console \> Show syntax highlighting in console input

I don't recommend rearranging your pane layout at this time.

## Running Code

Try running some code

### Load Library

```{r}
library(ggplot2)
```

### Some Scripting

```{r}
# This is a comment
# R runs line by line, or highlighted text

quiz1 <- 88 # how to assign values to names
quiz2 <- 94
quiz3 <- 79

quiz1

# averaging
(quiz1 + quiz2 +quiz3) / 3

mean(quiz1 + quiz2 + quiz3) # Not correct. Adds them together
mean(quiz1, quiz2, quiz3) # Not correct. Only returns mean of first argument.

mean(c(quiz1, quiz2, quiz3)) # Correct

list <- c(quiz1, quiz2, quiz3) # Assign multiple values to one list

mean(list)

average <- mean(list)

quiz3 <- 99 # Can update values

mean(c(quiz1, quiz2, quiz3))
list <- c(quiz1, quiz2, quiz3)

average <- mean(list)

average
```

### More Scripts

```{r}
scores <- c(88, 94, 79, 91, 85)
scores

mean(scores)
min(scores) # smallest number
max(scores) # largest number
length(scores) # size of list

summary(scores) # all those in one keyword

students <- c("a", "b", "c", "d", "e")
students

# bar plot

barplot(scores, names.arg = students)

# customized bar plot

barplot(scores, names.arg = students, ylim = c(0, 100), main = "TITLE", xlab = "pupils", y = "what they got")
```

## Introducing `ggplot2`

### Make a dataframe

```{r}
grades_df <- data.frame(student = students, score = scores)

grades_df

summary(grades_df)

nrow(grades_df) # number of rows
ncol(grades_df) # number of colums
dim(grades_df) # dimensions
```

### Try ggplot

```{r}

# make a foundation
ggplot(data = grades_df, aes(x = student, y = score))

# add layers
ggplot(data = grades_df, aes(x = reorder(student, -score), y = score, group = 1)) +
  geom_col(fill = "skyblue") +
  geom_point() +
  geom_line() +
  theme_minimal() +
  labs(title = "A GRAPH",
       subtitle = "QUIZ SCORES",
       x = "students",
       y = "scores")
```

### Mutate the dataframe

```{r}
library(tidyverse)

grades_df <- grades_df |>
  mutate(study_hours = c(2, 5, 1, 4, 3))
```

### Make a scatter plot

```{r}
ggplot(data = grades_df, aes(x = study_hours, y = score)) +
  geom_point() +
  geom_smooth(method = "lm", color = "skyblue") +
  theme_classic() +
  labs(title = "SCATTER",
       subtitle = "QUIZ SCORES",
       x = "Score",
       y = "Study Hours")

# cannot use fill (use color instead), don't use reorder

```

## Extra Credit (1pt)

What result do you expect when you run this expression?

```{r}
1 + 1 + 1

quiz1 + quiz2

1 + 1 + 1 - 3

0.1 + 0.1 + 0.1 - 0.3

print(0.1, 20)
print(0.3, 20)
print(0.5, 20)
```

What did you observe?

> Answer: Outputted numbers are not always exact, despite how they are depicted in code.

Why do you think this is happening? (You can look online for help)

> Answer: It's because certain numbers that aren't infinite floating points in base-10 are in base-2, or binary. 0.1 , 0.2, and 0.3 are infinite repeating decimals in base-2.
